### 2019-02-07 main.tex
Updates from Overleaf

\documentclass{llncs}
%\documentclass[converence,10pt]{IEEEtran}
%\IEEEoverridecommandlockouts
%\usepackage{cite}
%\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
%\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{bm}
%\usepackage{xcolor}
%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newcommand{\Exists}{\bm{\exists}\kern-0.6em\bm{\exists}}
\newcommand{\Forall}{\bm{\forall}\kern-0.6em\bm{\forall}}

\input{macros}

\begin{document}

\title{An Extension of LTL with Rules\\ 
and its Application to Runtime Verification}

\ignore{\author{
\begin{tabular}{ccc}
Klaus Havelund & \vspace{0.2cm} & 
Doron Peled \\
NASA Jet Propulsion Laboratory, Pasadena, USA & & 
Bar Ilan University, Ramat Gan, Israel \\
klaus.havelund@jpl.nasa.gov & & doron.peled@gmail.com
\end{tabular}
}
}

\ignore{
\author{\IEEEauthorblockN{Klaus Havelund} 
\IEEEauthorblockA{\textit{Laboratory for Reliable 
Software} \\
\textit{NASA Jet Propulsion Laboratory}\\
Pasadena, USA \\
klaus.havelund@jpl.nasa.gov}\\\mbox{}\\
\and
\IEEEauthorblockN{Doron Peled} 
\IEEEauthorblockA{\textit{Dep. of Computer Science} \\
\textit{Bar Ilan University}\\
Ramat Gan, Israel \\
doron.peled@gmail.com}
}
}

%\author{Submitted Anonymosly}

%\thanks{Copyright 2018. All rights reserved.}

\iffalse
\author{ 
Klaus Havelund~\thanks{The research performed by this author was carried out at Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration.} 
\and
Doron Peled~\thanks{The research performed by this author was partially funded by Israeli Science Foundation grant 1464/18: 
``Efficient Runtime Verification for Systems with Lots of Data and its Applications''.
}
}

\institute{
Jet Propulsion Laboratory, \\
California Institute of Technology, USA 
\and
Department of Computer Science \\
Bar Ilan University, Israel
}

\fi

%\hfill
\maketitle

\begin{abstract}

Runtime Verification (RV) consists of analyzing execution 
traces using formal techniques. 
%RV is scalable and 
%therefore has a big potential for infusion in software 
%engineering practices. 
It includes monitoring the 
execution of a system against properties formulated in 
Linear Temporal Logic (LTL). 
LTL offers a succinct notation for writing useful 
specification properties. However, it is limited in 
expressiveness in the propositional case, and several 
theoretic extensions have therefore been proposed. Furthermore, for many practical cases, there is a need to monitor properties 
that carry data, where one can use formalisms like 
first-order LTL. We show that 
first-order LTL has similar expressiveness
limitations as the propositional version. We suggest here 
two related extensions for increasing the expressive power: 
one for propositional LTL and one for 
first-order LTL. These extensions have a simple incremental operational 
semantics that is more suitable for RV than previously 
suggested extensions. We show that the 
propositional extension has the same expressiveness as the 
classical extensions, and demonstrate its
adoption for runtime verification of first-order safety 
properties. Finally, we expand the BDD-based runtime verification 
tool \dejavu{}\footnote{The actual tool name is not given due to anonymous reviewing.}  to support our extension and perform some 
experiments.

\end{abstract}

\input{rule-ltl}

\input{implementation}

\input{conclusion}

\input{prop.bib}

\newpage
\appendix

\section{Appendix}

%\klaus{In the text below mention what $i$ is (time point, position, ...}

\subsubsection{Construction of a property
for Theorem~\ref{theo2}.}
The model consists of
a temporal relation.
%\footnote{We abuse notation and
%use the same name for the relation and for its
%denotation in the logic. Also, $r [ i ]$ has zero
%parameters; as noted before, this is a natural %extension to the syntax and semantics presented %above.} 
$r$ with only the time parameter, so $r [ i ]$
has a Boolean value at the $i$th state. We will henceforth say that $r$ occurs
in state $i$. Another relation is $v$, with
two parameters. For each $i$, if $r$ does not appear in the
state, 
then there is a value $a$ from some (infinite) domain,
such that $v ( a, i )$, 
i.e., $v(a)$ appears in $i$.
The property requires that there are no two sets of values appearing within the relation $v$ (as $v(a)$) between successive occurrences of $r$ that are the same.

%\klaus{Below: isn't it the other way around, the first is not a model and the second is a model?}

Hence, the sequence of states 
\[ \{ r \} \{ v(1) \}  \{ v(2) \}  \{ v(2) \}  \{ r\}  \{ v(2) \}  \{  v(1) \}  \{  v(5) \}  \{  r \}\]
is a model of this formula, while 
\[ \{ r \} \{ v(1) \}  \{ v(2) \} \{ v(3) \}  \{ r\}  \{ v(2) \}  \{  v(1) \}  \{  v(3) \} \{  r \} \]
is not.
The QFLTL formula is a bit involved, hence given in several parts (macros), where the main formula appears last.

\begin{itemize}
\item $add(P) ::= \forall x \, (P(x) \leftrightarrow (v(x) \vee \ominus P(x))$ [The set $P$ differs in the current state from the previous one only by values $a$ that appear in $v(a)$.]
\item $same(P) ::= \forall x \, (P(x) \leftrightarrow \ominus P(x))$ 
[The set of elements $x$ for which
$P (x)$ holds
contains the same elements as in the previous state.]
\item $empty (P) ::= \forall x \neg P(x)$ [The set $P$ is empty.]
\item $equal (P,Q) ::= \forall x (P(x) \leftrightarrow Q(x))$ 
[The sets represented by the relations
$P$ and $Q$ contain the same elements {\em in the current state}.]
\item $collect(P, \varphi, \psi) ::=
       r \wedge same(P) \wedge \varphi \wedge 
       \ominus ((add(P)\wedge \neg r \wedge \varphi) \ \since \ (r \wedge empty(P) \wedge \psi))$  [Collect elements $a$ in $v(a)$, while $\varphi$ holds, 
       since the previous state where $r$ 
       and $\psi$ held.]
\item $init ::= r \wedge \Box (r \leftrightarrow \neg \exists x \, v(x))$ [$r$ holds in the first state, and $v(a)$ with some value $a$ holds in states where $r$ does not hold.]
\item $init \wedge \Box \neg \Exists A \Exists B (equal (A, B) \wedge \\
collect(B, same(A),\, same(A)) \ \since \ collect(A, true, true))$
[There are no two sequences of states where $v(a)$ holds
with various values $a$, between a pair of states satisfying $r$ that contain exactly the same elements (regardless of repetition).]
\end{itemize}

%\klaus{Above: maybe need to point out what the top formula is.}

%\klaus{Above: isn't there missing a $\Box$ in the main formula?}

\iffalse
We prove that the above property cannot be expressed in EPFLTL.
We use a combinatorial argument, showing that
the information that is available in any EPFLTL
property
after a prefix is not sufficient to distinguish between two models (executions), one satisfying and one not satisfying it.

In EPFLTL, after every prefix of the input, it is
sufficient to keep a finite set of (non temporal) relations over the
data elements seen in the input, which
summarizes the prefix. This follows from the set
semantics, which appears in Section~\ref{EPLTLRV}, 
where the interpretation of
each subformula, and the auxiliary relations,
at a particular state is a relation, and the interpretation 
of a property a prefix of length $i$ only depends on
these relations. Now, the amount of information
that summarizes an observed prefix for
EPFLTL is is polynomial
in $N$, consisting of a finite number
of relations over
$N$ elements. The number of different tuples
for such relations can be shown, inductively, to
be polynomial in $N$ (beach parameter
of each relation can be one out of the $N$ possibilities, or a representation the values not seen yet).
However, the above property depends on whether
a set of values that appear as parameters to
the relation $v$, between two consecutive appearances
of $r$'s, are the same. For that, one needs to keep some representation of all the
subsets of the elements that appeared in the input so far; by contradiction, any summary representation of the observed prefix that
would identify two prefixes with some different sets, would not be able to distinguish between
sequences that extend this prefix with the set
of elements that appears in one prefix but not the other.
But the number of such subsets is exponential in $N$. \qed
\fi

%\klaus{Above: why set of subsets?}

\iffalse
{\bf Comments.} First, note that the reason that QFLTL {\em can} express the property in Theorem~\ref{theo2}, is that it allows to augment the input relations with auxiliary relations that are not fixed. That is, for an auxiliary relation $R$,
$R [ i ]$ is not fixed, and can change per different
prefixes (of length at least $i$), or even in different
subformulas. Second, given a simple extension to EFLTL, where 
one can use a variable that represents an updatable
counter (or a generator of distinct values), one can express the property used in~\ref{theo2}, by
keeping a relation where values between different occurrences of the event $r$ are paired with distinct values, hence, representing an unbounded set of sets.
\fi

%%%M
\subsection*{Details of left to right direction of Theorem~\ref{theo1}.}
One can translate any QLTL property into a second order monadic logic in a standard
way~\cite{Thomas}, and then from
second order monadic logic to a deterministic Muller automaton $M = (S, \iota , \Sigma, \Delta, {\cal F})$ (a B\"{u}chi automaton may not always be deterministic), where $S$ is the set of states, $\iota \in S$ is the initial state, $\Sigma$ is the set of inputs, each $\alpha \in \Sigma$ is a minterm (i.e., conjunction of
each variable in the set negated or non-negated) of the variables in $B$, $\Delta : S \times \Sigma \mapsto S$ is the transition function, and ${\cal F} = \{ F_1 , \ldots , F_k \}$ is the accepting set. A muller automaton 
accepts a run if it 
passes infinitely often in exactly
one of the specified accepting sets. To formulate an equivalent ELTL formula,
the auxiliary variables represent the states $S$ of $M$, i.e., $a_s$ represents the state $s \in S$.

%\klaus{The following formulation above is not not clear: ``where an accepted input needs to pass infinitely often in exactly
%one of the specified sets''.}

\noindent
Then we translate $M$ into an ELTL formula
of the form:

\begin{tabbing}
\ \ \ \ \ $\bigvee_{F \in {\cal F}} \bigwedge_{s \in F} ( \Box\Diamond a_s \wedge \bigwedge_{s \not\in F} \Diamond \Box \neg a_s)$
where \\
\ \ \ \ \ \ \ \ \ \ $a_s :=   ( \alpha_{\iota} \wedge \ominus \false ) \vee \bigvee_{\Delta ( s'
, \alpha ) = s } 
(\alpha \wedge \ominus a_{s'}) 
\, :\,  s \in S$
\end{tabbing}

%\klaus{Above: should $s=i \rightarrow \neg \ominus \true$ instead be (note: two changes):
%$\neg \ominus \true \rightarrow a_s=i$
%}
%\klaus{Above: can't judge whether the $\bigwedge_{s \not\in F} \Diamond \Box %\neg a_s)$ is correct.}
\noindent
Basically, the assignments to the auxiliary variables are used to simulate the change of states\footnote{It is possible to use only $\lceil \log ( | S | ) \rceil$
variables, at the expense of longer and more complicated formulas.}, and the main formula expresses the Muller acceptance condition. \qed

\subsubsection*{Set Semantics.}

Let $\Gamma$ be
a set of assignments over the variables $W$, and
$U \subseteq W$.
Then $\proj ( \Gamma , U )$ (for ``projecting {\em out}'' or  ``hiding''
the variables $U$)
is the largest set of assignments over 
$W \setminus U$, 
each agreeing with some
assignment of $\Gamma$ on all the variables in $W \setminus U$.
Let $U \cap W = \emptyset$, then
$\ext ( \Gamma , U )$ is the largest set of assignments
over $W \cup U$, where each such assignment agrees 
with some assignment in $\Gamma$ on the
values assigned to the variables $W$. 
\iffalse
This means that we extend $\Gamma$ by
adding arbitrary values to the variables in $U$ from their domains.
Then $\proj ( \ext ( \Gamma , U ) , U ) = \Gamma$ holds.
We define the union and intersection operators on sets of
assignments, even if they are defined over non identical
sets of variables. 
In this case, the assignments are extended
over the union of the variables. 
\fi
Thus, if $\Gamma$ is a
set of assignments over $W$ and $\Gamma'$ is
a set of assignments over $W'$, then
$\Gamma \; \bigcup \;  \Gamma'$ is defined as $\ext (\Gamma , W' \setminus W ) \cup
\ext ( \Gamma' , W \setminus W' )$ and
$\Gamma \; \bigcap \; \Gamma'$ is $\ext (\Gamma , W' \setminus W ) \cap
\ext ( \Gamma' , W \setminus W' )$.  Hence, both are defined
over the set of variables $W \cup W'$.

We denote by $A_{\vars ( \varphi )}$ the set of all possible assignments
of values to the variables that appear free
in $\varphi$. Thus,
$I [ \varphi, {\cal R} , i] \subseteq A_{\vars ( \varphi )}$.
To simplify definitions, we add
a dummy position $0$ for sequence $\sigma$ (which starts with $s_1$), 
where every formula is interpreted as an empty set.
Observe that the value $\emptyset$ and $\{ \epsilon \}$, behave
as the Boolean constants $0$ and $1$, respectively.
The set semantics is defined as follows, where $i \ge 1$.

%\klaus{Above: assignments are introduced without any gentle introduction. Remember: you are not communicating with Moshe Vardi. You are communicating with John Fingelstein who had three glasses of wine at dinner.}

%\klaus{Above: $I [ \varphi, {\cal R} , i]$ is introduced out of the blue without any explanation.}

\begin{itemize}
\item $I [ \varphi , {\cal R} , 0 ] = \emptyset$.
\item $I [ \true , {\cal R} , i ] = \{ \epsilon \}$.
\item $I [ p ( a ) , {\cal R} , i ] =$ if $m (p )  ( a , i ) $ then
$\{ \epsilon \}$ else $\emptyset$.
\item $I [ p ( x ) , {\cal R} , i ] = \{ [ x \mapsto a ] | m (p) ( a , i)  \}$.
%\item $I [ \true , \sigma , i ] = D^n$.
%\item $I [ ( \varphi \vee \psi ) , \sigma , i ] = 
%I [ \varphi , \sigma , i ] \cup I [ \psi , \sigma , i ]$.
\item $I [ ( \varphi \wedge \psi ) , \sigma , i ] = 
I [ \varphi , {\cal R} , i ] \;  \bigcap \; I [ \psi , \sigma , i ]$.
\item $I [ \neg \varphi , {\cal R} , i ] = 
A_{\vars ( \varphi )} \; \setminus \; I [ \varphi , {\cal R} , i ]$.
\item $I [ ( \varphi \; \since \; \psi ) , {\cal R} , i ] = 
I [ \psi , {\cal R}, i ] \; \bigcup \;
( I [ \varphi , {\cal R} , i ] \bigcap 
I [ (\varphi \since \psi ) , {\cal R} , i - 1 ] )$.
\item $I [ \ominus \varphi , {\cal R} , i ] = I [ \varphi , {\cal R} , i-1 ]$.
\item $I [ \exists x \; \varphi , \sigma , i ] = 
\proj ( I [ \varphi , {\cal R} , i ] , \{ x \} )$.
\item $I [ \psi \mathrm{\ where\ } r_j ( x_j) := 
\varphi_j (x_j) : j \in \{ 1 , \ldots , n \} , {\cal R} , i ] = 
 I [ \psi , {\cal Q} ] $ where ${\cal Q} = {\cal R} \cup
\{ r_j | r_j [ i ] =  I [ 
\varphi_j , {\cal Q} , i ] \}$
\end{itemize}
Note that the last time interprets a formula of
the form defined in~(\ref{EFLTL}).

%\klaus{Above: as the semantics is presented, it looks as if 
%the `$\psi\ where ...$' is just another subformula.}

%\klaus{Above: hide is not defined.}

%\klaus{Above: the semantics of $p(a)$ and $p(x)$ do not refer to 
%${\cal R} $.}

In the last item, the extended model ${\cal Q}$ appears both on the left and the right of the equality sign. It is well defined because of Lemma~\ref{samesame}.
\end{document}