### 2019-03-27 RESPONSE.txt
added response

General comment: Two of the reviews seem to ask (more or less
directly) what the contribution of the paper is.  We admit that the
introduction may not have provided a sufficiently clear overview over
this work, since many logics are introduced, but in the following we
try to briefly outline this (and we will of course later improve the
introduction).

In previous work [23,24] we presented a BDD-based monitoring algorithm
for:

  PFLTL : Past-time First-order LTL

as defined on page 9. This logic has limitations in expressiveness as
shown with two examples on page 10. As main contribution, we
consequently present an extension of this logic, called:

  EPFLTL: Extended Past-time First-ordrer LTL

See page 11 equation (5). The extension coniststs of adding rules to
temporal properties.  Example properties (not expresssible in PFLTL)
are shown in Figure 2 page 16.  This logic is strictly more expressive
than PFLTL by allowing e.g. Wolper's in-every-even-state property and
generally transitive closures over data in the trace to be
expressed. The algorithm is presented on page 14.  The corresponding
tool has been extended as well (and is availbale open source).

The paper is structured in such a way that it gradually builds up this
result in steps, with some intermediate results, first making the
argument for the propositional LTL case in Sections 2 and 3, and then
for the first-order LTL case in Sections 4 and 5, as follows.

For the propositional case:

- LTL is introduced and its limitation to past time operartors PLTL
  (known results)

- QLTL (and its past time version QPLTL) is LTL (PLTL) with (so-called
  dynamic) quantification over states. This is a known extension
  described by Wopler in [21]. This extends the expressiveness of LTL
  to allow formulation of the property: "every even state satisfies
  p", not expressible in LTL. However it is not obvious how to design
  an algorithm for efficiently monitoring QLTL properties. Therefore
  we alternatively propose (and this is the main theme in the paper):

- ELTL (and its past time version EPLTL) that instead extends LTL
  (PLTL) with rules, achieving the same expressivness as QLTL (QPLTL)
  but with a well defined monitoring algorithm. This is a new result.

For the first-order case we repeat the above story, ending with the final
contribution:

- FLTL and specifically PFLTL are introduced. The logic PFLTL and its
  monitoring algorithm was introduced in previous work by us [23,24].

- QFLTL (and its past time version QPFLTL) is introduced, showing how
  to extend FLTL with quantification over states as in the
  propositional case. This is a new theoretic result. However, as in
  the propositional case, it is not clear how to construct an
  efficient monitoring algorithm. This therefore leads us to the final
  result and main contribution:

- EFLTL and specifically the past time version EPFLTL are introduced,
  and a monitoring algorithm for EPFLTL is presented.

----------------------- REVIEW 1 ---------------------

> The authors claim that these extensions are useful for runtime
> verification and present some minimal experimental results from
> running a BDD-based implementation of the extensions.

As for experiments with BDDs, we have performed many experiments, they
are mentioned in our past papers, and are publically available with
our tool. But due to anonymous reviewing, we cannot reveal it. We only
here presented experiments to show the strength of the extension of
the logic.

> I am not quite on board with the motivation of the paper. In runtime
> verification, the truth values of formulas are determined on a single
> trace, hence, in particular, LTL and CTL are equivalent, and the
> verification is hence polynomial. It is also quite standard in
> model-checking to augment the intermediate states on the path with the
> truth values of subformulas (this is how CTL model checking is done,
> and hence also LTL for a single trace). Wouldn't it address at least
> some of the concerns of the authors (that is, which subformulas are
> satisfied in which states)?

*** TODO ***

We do not quite understand this comment.

> Also, I see no reference to PSL, which is the IEEE standard for
> property specification languages, or to ForSpec, which is Intel's
> specification language. I am wondering whether these would already
> provide the necessary functionality.

We should definitely cite the excellent well known work on
PSL/ForSpec. However, PSL and other standards are limited in the way
they handle data, and are basically strong propositional logics. They
are equivalent to logics that we cited, e.g. regular expressions and
2nd order monadic logic (this is not a logic with data!). We
introduced the propositional extension with rules (ELTL/EPLTL) to
demonstrate how simple (not more expressive than regular expressions)
our extension is, and as an easy to read motivation for a related
extension for first-order LTL (EFLTL/EPFLTL).  The subject of our
paper is not about expressiveness of propositional LTL, but a
practical extension to first-order LTL. As a side effect, the
propositional introduction also makes a new nice style of expressive
LTL properties.

> Adding quantification is solvable, in general, with a QBF solver, and
> in some cases even without it. In the Intel's paper about enhanced
> vacuity from the beginning of 2000's, they studied the problem of
> model-checking a quantified formula and demonstrated that for the
> types of formulas they constructed it is no harder than
> model-checking.

We assume you are referring to the paper:

  Enhanced Vacuity Detection in Linear Temporal Logic
  R. Armoni, L. Fix, A. Flaisher, O. Grumberg, N. Piterman, A. Tiemeyer, M. Vardi
  CAV 2003

The main difference is the first-order nature of our logic, in
contrast to the logic presented in that paper, which is propositional
(and by the way limited in where quantifiers can appear). However,
thanks for the reference, we will explore this work.

> Finally, I might be wrong, but it seems to me that BDDs are no longer
> widely used in model-checking, replaced by SAT and SMT solvers. I am
> wondering whether the implementation with BDDs is interesting enough.

The main problem studied in this paper is that of a linear temporal
logic with data, i.e., (past) first order linear temporal logic. For
this reason, a representation that can efficiently handle relations
over (basically unbounded) data is important. The BDD representation
is used to represent such big relations, where the BDDs hold
enumerations of the seen data values rather than that (hashed) data
values themselves. Consequently, we can use efficient standard Boolean
operators on BDDs to implement intersection, union and complementation
of sets (since the sets are only semi infinite, there is a special
encoding value for "all values not seen"). This is a surprisingly
efficient representation which works with first-order LTL.  BDDs are
nowdays less frequently used for model checking, but (a) this is not
model checking but runtime verification (b) they are used for the
specific and ingenious way of representing relations, and (c) SAT
solvers are not needed here. We are aware of (and will cite) work on
using SMT for runtime verification:

  Monitoring Modulo Theories
  Normann Decker, Martin Leucker, and Daniel Thoma

but it is not particularly efficient.

----------------------- REVIEW 2 ---------------------

> While the paper explores a potentially useful abstraction for runtime
> verification, it was hard to fully understand the motivation for this
> work. Some ways that the work could be strengthened are listed below.

Sure, see also leading comment above.

> Thm. 1 and Lemma 2 show that the introduced logics (ELTL and EPLTL)
> have the same expressive power as Q(P)LTL. The authors should make the
> benefit of introducing these new logics more clear.
> 
> Thm. 2 shows that
> QPFLTL is more expressive than EPFLTL, but the authors do not provide
> any complexity results on the decidability of these logics - for what
> benefit are we trading the expressivity?

As explained in the general introduction of this response, the main
reason for moving from 'Q' to 'E', e.g. from QPFLTL to EPFLTL, is that
we can show how to construct an efficient monitoring algorithm for 'E'
(e.g. EPFLTL), whereas we do not see how to do this for more theoretic 'Q'
(e.g. QPFLTL). That is the main message of the paper.

> Thm 2 also opens the question
> of whether EQPFLTL makes sense and if so how it relates to the other
permutations of (E/Q/F/PLTL).

There is no need to extend QPFLTL with rules since the rule extension 'E'
is an easier monitorable *alternative* to the dynamic quantification 'Q'.

> The limitations of existing runtime verification systems for data
> intensive systems could be made more clear through an evaluation that
> compares the use of EFPLTL for runtime verification with the
> alternative existing approaches. At the moment the evaluation only
> presents benchmarks for the introduced tool, but does not situate this
> evaluation in the context of existing runtime verification tools.

That is a fair comment.

> Furthermore, the evaluation includes a 'spawning' trace where analysis
> time appears to scale exponentially in the size of the trace. The
> authors mention this is possibly due to the increasingly large amount
> of data that must be stored as the trace grows. As the system is
> intended to be used for runtime verification, what practical
> consideration must be made given the exponentially scaling with
> respect to the lifecycle of the task being monitored?

It is certainly true that dealing with data introduces the problem
that monitoring becomes trace length dependent.  To tackle this
problem, the original version of the tool performs a clever form of
garbage collection, described in [24], where data no longer needed are
dropped from the BDDs.  In the here (artificially created) property
and trace, there is indeed a problem if too many threads are spawned.
We have not in this work studied how to limit monitoring to become
less trace length dependent. We consider this as an orthogonal,
although very important, line of work in the field of runtime
verification. Only little work has been done in this direction so far.

> In addition, a more thorough analysis of memory usage should be
> included in the evaluation. No information is provided on the system
> that ran these evaluation results. Do these benchmarks' memory usage
> exceed the available RAM and trigger disk operations? As the temporal
> relations of (P)FLTL require increasingly large hash tables to save
> their evaluations, what impact will this have on memory limited
> systems? This is especially relevant for the many use cases of runtime
> verification on embedded systems.

Indeed, memory analysis should be included. This information was not
directly visible when running our scripts (not a good excuse,
admitted). Memory is of course critical. As mentioned above, we do
perform garbage collection on unused values as described in [24]. For
example for the property

  forall x . p(x) -> !q(x) S r(x)

(meaning: forall x . if p(x) then in the past r(x) and since then no
q(x)), the monitor will drop any value x for which we in the past saw
an r(x) and since then saw! a q(x). But the general problem of
monitoring embedded systems is challenging. Not all problems can be
solved at once.

> Another issue is that it was hard to identify the paper's
> contribution. The section on 'Runtime verification algorithm for
> PFLTL' has already been described in [24]. To handle EPFLTL, Sec 5
> introduces a single 'rename' function which seems a relatively small
> delta with respect to [24]. The authors point us back to section 3
> (mixed evaluation order) for further contributions with respect to
> [24] to handle EPFLTL, but it is then not clear why Section 3 is
> separate from Section 5.

See general introduction to this response for a more general outline
of the work. The handling of EPFLTL, compared to PFLTL, is not trivial
and was quite difficult to implement. We do not consider it a
relatively small delta, although the theory perhaps is small (and
elegant). Concering the question why Section 3 is separate from
Section 5: as outlined in the introductory text above, Section 3 deals
with the propositional case whereas Section 5 deals with the
first-order case.

> The paper should better position its contributions wrt. related work.
> Throughout the paper, the authors should make it more obvious to the
> reader that only logics prefixed with an E are new logics introduced
> in this work (if I understood the paper correctly).

Well, QFLTL and QPFLTL are new (first-order LTL extended with dynamic
quantification).  But yes, the comment is largely correct. Our
introduction did not clarify the structure well enough.

> Additionally, the use of predicates to encode information about the
> prefix of the state (system inputs seen so far), and thereby allowing
> arbitrary data over infinite domains, was explored in 'Synthesizing
> Functional Reactive Programs' by Finkbeiner et. al
> https://arxiv.org/abs/1712.00246.

*** TODO ***

> Finally, the paper contains typos that create problematic ambiguities.
> As an example, the top of page 5 notes that the logic PQLTL is
> obtained by restricting QPLTL to past modalities. From context, it
> appears this should read as QLTL, not QPLTL.

Indeed.

> Furthermore, by page 6
> the authors are proving properties of QPLTL, but PQLTL is no longer
> referenced. Although from context the reader can infer that QPLTL and
> PQLTL these should be taken as the same logic (except on page 5 when
> QPLTL refers to QLTL), these inferences should not be left to the
> reader. I also believe that there is an issue with Theorem 2 - when
> comparing the expressive power of EPFLTL and QFPLTL, the statement
> instead reads as comparing EPFLTL and QFLTL.

Indeed to all observations here, we should have used/should use latex
macros for this!

----------------------- REVIEW 3 ---------------------

> The experimental results seem
> preliminary: apart from the fact that the examples are made up, it's
> unclear how the collection of traces was generated.

The artificial traces were generated by a trace generating program
parameterized with various constants, indicating how many events of
various kinds should be generated, and the patterns with which they
are generated. The objectieve was to generate as many data values in
the trace as possible, while at the same time also cause repetition of
patterns.

> It would be more
> convincing to see this applied to more realistic examples (esp the
> transitive closure property surely has many applications).  However,
> the tool is apparently available publicly (links hidden due to blind
> review).

Yes, it would be better with more realistic examples. The tool is
publicly available.

------------------------------------------------------