### 2019-07-15 abstract.tex

\begin{abstract}

Runtime Verification (RV) consists of analyzing execution 
traces using formal techniques, e.g., monitoring executions against 
Linear Temporal Logic (LTL) properties. Propositional LTL is, however, 
limited in expressiveness, as first shown by Wolper \cite{Wolper}.
%, e.g. not express that every even state in a sequence satisfies a given proposition. 
Several extensions to propositional LTL, which promote the
expressive power to that of regular expressions,
have therefore been proposed; however, none of which was, by and large, adopted for RV.
In addition, for many practical cases, there is a need in RV to monitor properties 
that carry data. This problem has been addressed by numerous authors, and 
in previous work we addressed this by providing an algorithm that uses BDDs to represent relations over data elements.
We show expressiveness deficiencies of first-order LTL and suggest an extension of (propositional as well as first-order) LTL with rules to address these 
limitations.  We describe how the \dejavu{} tool is correspondingly extended and provide some experimental results. 
%It has an incremental operational 
%semantics that suggests a simple extension for RV algorithms. 
%We demonstrate that by expanding the \dejavu{} tool to 
%support our extension, and performing some experiments.

\end{abstract}