### 2019-07-15 rule-ltl.tex

\newcommand\eventty{\mathbb{E}}
\newcommand\setof[1]{\mathcal{P}(#1)}

\section{Propositional LTL}
\label{propLTL}

The classical definition of linear temporal logic~\cite{MP} has the following syntax:
\[ \varphi ::= \true \, | \,  p \, | \, ( \varphi \wedge \varphi ) \, |  \, \neg  \varphi \, |   \, \bigcirc \varphi \, |  
\, ( \varphi  \until  \varphi ) \,|  \, \ominus \varphi |
   \, ( \varphi \ \since \ \psi ) \,  \]
where $p$ is a proposition from a finite set of propositions $P$, and $\bigcirc$, $\until$, $\ominus$, $\since$ stand for {\em next-time}, 
{\em until}, {\em previous-time} and {\em since}, respectively.
%
The models for LTL formulas 
are infinite sequence of states, of the form
$\sigma =  s_{1}\, s_{2}\, s_3 \ldots$,
where $s_i \subseteq P$ for each $i\geq 1$. These are the propositions that {\em hold} in that state.
%We denote by $\xi_{i}$ the suffix $s_{i} s_{i+1} s_{i+2}\ldots$ of $\sigma$. 
LTL's semantics is defined as follows:
\begin{itemize}
\item $( \sigma, {i}) \models \mathit{true}$. \ \ \ \ \ 
\item $(\sigma, {i}) \models p$ iff $p \in s_{i}$. \ 
\item $(\sigma, {i}) \models \neg\varphi$ iff  $(\sigma, {i}) \not\models \varphi$.
\item $(\sigma, {i}) \models ( \varphi \land \psi)$ iff $(\sigma, {i}) \models \varphi$ and $(\sigma, {i} ) \models \psi$.
\item $(\sigma, {i}) \models \bigcirc \varphi$ iff $(\sigma, {i+1}) \models \varphi$.
\item $(\sigma, {i}) \models ( \varphi  \cal{U}  \psi )$ iff for some $j$,
$j\geq i$, $(\sigma, {j}) \models \psi$,
and for each $k$, $i \leq k < j$, $(\sigma, {k}) \models \varphi$.
\item $(\sigma , i) \models \ominus \varphi$ iff $i > 1$ and $(\sigma, i-1) \models \varphi$.
\item $(\sigma , i) \models (\varphi \since \psi)$ iff there exists $j$, $1 \leq j \leq i$, such that
$(\sigma , j)  \models \psi$ and for each
$k$, $j < k \leq i$, $(\sigma , k ) \models \varphi$.
\end{itemize}
Then $\sigma \models \varphi$ when $( \sigma ,1 ) \models \varphi$.
We can use the following abbreviations:
$\false = \neg \true$, 
$(\varphi \vee \psi) = \neg ( \neg \varphi \wedge \neg \psi )$, 
$(\varphi \rightarrow \psi) = ( \neg \varphi \vee \psi)$,
$(\varphi \leftrightarrow \psi) = ((\neg \varphi \wedge \neg \psi) \vee (\varphi \wedge \psi ))$,
$\Diamond \varphi = (\true \until \varphi)$, $\Box \varphi = \neg \Diamond \neg \varphi$,
${\bf P}\  \varphi = (\true \; \since \; \varphi)$ ({\bf P} stands for P{\em reviously}) and 
${\bf H}\  \varphi = \neg {\bf P}\ \neg \varphi$ ({\bf H} stands for 
H{\em istory}).

%%infinite sequences (finite executions can be extended by repeating the last event indefinitely to form an infinite paths).
The expressive power of different versions of propositional LTL is often compared to regular expressions
over the alphabet $\Sigma = 2^P$ and to  {\em monadic} first and second-order logic.
%, where the  propositions in the LTL formula become
%unary predicates over the naturals, and with the $<$
%relation. 
 Accordingly, we have the following characterizations:
LTL is equivalent to monadic first-order logic,  star-free regular expressions\footnote{Regular expressions without the star operator (or $\omega$).} and counter-free B\"{u}chi automata.
%\footnote{A counter-free language is a regular language for which there is an integer $n$ such that for all words x, y, z and integers $m \geq n$ we have that $xy^m z \in L$ if and only if $xy^n z \in L$.}. 
For an overview of logic and automata see~\cite{Thomas}.
Restricting the temporal operators to the {\em future}  operators $\until$ and $\bigcirc$
(and the ones derived from them $\Box$ and $\Diamond$)
maintains the same expressive power.

An important subset of LTL, called here \PLTL{}, allows only past temporal
operators: $\since$, $\ominus$ and the operators derived
from them, $\mathbf{H}$ and $\mathbf{P}$. The past-time logic is sometimes interpreted over finite sequences, 
where $\sigma \models \varphi$ when $( \sigma , | \sigma | ) \models \varphi$.
It is also a common practice to use a \PLTL{} formula, prefixed with  a single $\Box$ (always) operator; in this case, {\em each of the prefixes} has to satisfy $\varphi$. This later
form expresses {\em safety} LTL properties~\cite{AS}.
When \PLTL{} is interpreted over finite sequences, its
expressive power is the same as star-free regular expressions, first-order monadic logic over finite sequences and counting-free automata.

%\section*{Extending Propositional LTL}
\label{sec:extending-prop-ltl}

Wolper~\cite{Wolper} demonstrated that the expressive power of LTL is lacking using
the property that all the states with even\footnote{This is
different than stating that $p$ alternates between $\true$ and $\false$ on consecutive states.} indexes in a sequence
satisfy some proposition $p$.

%Wolper's example property cannot be expressed in \PLTL{} either.
%In order to extend the expressiveness of LTL, Wolper suggested to use linear grammars, or, alternatively, the ability to quantify over
%propositions, as shown below.

\vspace{1.2ex}
\noindent {\bf Extending LTL with dynamic quantification.}
Adding quantification over propositions, suggested by Sistla in~\cite{sistla-phd-83}, allows writing a formula of the form
$\Exists q \, \varphi$, where
$\Exists q$ represents {\em existential} quantification over a proposition $q$ that can appear in $\varphi$.
%
%A model for such a formula 
%has truth assignments only over the propositions in $B$, but the propositions $A$ can appear (only) within the scope of a quantifier.
To define the semantics, let $X \subseteq P$ 
%be a subset of 
%the propositional variables appearing in the %states of the sequence $\sigma = s_1 s_2 %\ldots$
and denote $\sigma |_X = s_1 \setminus X \, s_2 \setminus X \ldots$. (Note that $\sigma |_X$ denotes projecting {\em out} the propositions in $X$.) The semantics is defined as follows:

\begin{itemize}
\item $( \sigma, i ) \models \boldsymbol{\Exists} q \, \varphi$ iff there exists
$\sigma'$ such that $\sigma' |_{\{ q \} } = \sigma$ and 
$( \sigma', i) \models \varphi$.
\end{itemize}
%Since all the auxiliary propositions appear in the scope of
%quantifiers, it holds that $\sigma \models \varphi$
%(in $\varphi$, auxiliary variables appear only within
%the scope of quantifiers)
%iff $\sigma |_A \models \varphi$.
%
%\klaus{Are the last two lines above needed?}
%
Universal quantification is also allowed, where $\Forall q\, \varphi =
\neg \Exists q\, \neg \varphi$.
This kind of quantification is considered to be {\em dynamic}, since the quantified propositions can have different truth values depending on the states. It is also called  {\em second-order} quantification, since the quantification establishes the {\em set} of states in which a proposition has the value $\true$.
Extending LTL with such quantification, the logic \QLTL{} has the same expressive power as regular expressions, full B\"{u}chi
automata, or monadic second-order logic with unary predicates over the naturals (see again~\cite{Thomas}).
In fact, it is sufficient to restrict the quantification to existential quantifiers that prefix the formula to obtain the full expressiveness of \QLTL{}~\cite{Thomas}. 
\iffalse
Wolper's property can be expressed in \QLTL{} as follows:
\begin{equation} \label{form0} \Exists q \,
( (  \neg q \wedge \Box (q \leftrightarrow \bigcirc \neg q )  )  \wedge \Box (q \rightarrow p ) )\end{equation}
\fi
Restricting \QLTL{} to the past modalities, one obtains the logic \QPLTL{}. \QPLTL{} has the same expressive
power as regular expressions and finite automata.
Wolper's property can
be rewritten in \QPLTL{} as:
\begin{equation} \label{form1} \Exists q \,
{\bf H} (( q \leftrightarrow \ominus \neg q )   \wedge (q \rightarrow p ) )\end{equation}
%In fact, it can be shown that a single existential quantifier is sufficient. 
Since $\ominus \varphi $ is interpreted as  $\false$ in the first state of any sequence, regardless of $\varphi$, then
$q$ is $\false$ in the first state. Then $q$ alternates between even and odd states.

%Note that since finite automata over finite sequences can be determinized, it is always sufficient to translate a
%\QPLTL{} property into a deterministic
%automaton.

\vspace{1.5ex}
\noindent {\bf Extending LTL with rules.}
We introduce another extension of LTL, which we call \ELTL{}.
%It has a form that puts it, conceptually, in between temporal
%logic and automata: on the one hand, it is possible to simply write
%a pure LTL formula, and on the other hand, one can write a specification that resembles a finite automaton. 
As will be showed later, this extension
is very natural for runtime verification.
%
We partition the propositions $P$ into
{\em auxiliary propositions} $A = \{ a_1 , \; \ldots\; , a_n \}$
and {\em basic propositions} $B$.
%A model for an \ELTL{} formula has truth assignments only over the propositions in $B$.
An \ELTL{} property $\eta$ has the following form: 
\begin{equation} \label{ELTL}
\psi \mathit{\ where\ } a_j  := \varphi_j \, : \, 
 {j \in \{1, \ldots , n\}} \end{equation}
where each $a_j$ is a distinct auxiliary proposition from $A$,
$\psi$ is an LTL property and each $\varphi_i$ 
is a \PLTL{} property where propositions from $A$ can
only occur within the scope of a $\ominus$ operator.
The semantics can be defined as follows. We refer to $\psi$
as the {\em statement} of $\eta$ and to 
$a_j  := \varphi_j$ as a {\em rule} (in text, rules
 will be separated  by commas).

\begin{center}
$\sigma \models \eta$ if there exists $\sigma'$, where
$\sigma' |_A = \sigma$ s.t.  
 $\sigma' \models ( \psi  \wedge   \Box \bigwedge_{1 \leq j \leq n} ( a_j \leftrightarrow \varphi_j)) $
 \end{center}

\noindent
Wolper's example can be written in \ELTL{}
as follows:
\begin{equation} 
\label{form3}
\Box ( q \rightarrow p ) \, \mathrm{where} \, 
  q := \ominus \neg q
\end{equation}
where $A = \{ q\}$ and $B = \{ p \}$.
The auxiliary proposition $q$ is used to augment the input sequence such that each {\em odd} state will satisfy $\neg q$
and each {\em even} state will satisfy $p$.  
%
\ELTL{} extends the set of propositions with new propositions, whose values at a state are
functions of (i.e., uniquely defined by) the prefix of the model up to that state. The added propositions summarize the past execution prefix up to and
including the current state.
This differs from 
the use of auxiliary propositions in \QLTL{}, where the values assigned to the
auxiliary propositions do not have to extend the 
states of the model in a unique way throughout the interpretation of 
the property over a model. The constraint that auxiliary propositions appearing in the formulas $\varphi_i$ 
must occur within the scope of a $\ominus$ operator
is required to prevent conflicting rules, as in
$a_1 := \neg a_2$ and $a_2 := a_1$.
%, which makes a formula
%that uses $a_1$ and $a_2$ uniformly $\false$. Still, one can write unsatisfiable rules, e.g., by using the rule $a_1 : = \false$.

\begin{lemma}[Well foundedness of auxiliary propositions] \label{fourone}
The values of the
auxiliary propositions
of an \ELTL{} formula $\eta$
are uniquely defined in a state of an execution
by the prefix of the execution upto and including that state.

\end{lemma}

%\klaus{(1) is this lemma saying the right thing? (2) is the proof really a proof of that lemma, (3) 
%$\varphi$s and the $\psi$s. Do we need a special greek symbol for it?}

\noindent {\bf Proof.}
Let $\eta$ be a formula over auxiliary propositions $A$ and basic propositions $B$,
with rules
$a_j  := \varphi_j \, : \, 
 {j \in \{1, \ldots , n\}}$.
%, with
%a set of rules $a_j  := \varphi_j \, : \, 
% {j \in \{1, \ldots , n\}}$. 
Let $\sigma$ be a model with states over $B$. 
Then there is a unique
model $\sigma'$ such that $\sigma' |_A = \sigma$ and $\sigma' \models \Box \bigwedge_{1 \leq j \leq n} ( a_j \leftrightarrow \varphi_j) $: inductively,
the value of each auxiliary proposition $a_j$ at the $i$th state
of $\sigma'$
is defined, via a rule $ a_j := \varphi_j$, where $\varphi_j$ is a \PLTL{} formula; 
hence it depends on
the values of the propositions $B$
in the $i$th state of $\sigma$, and on the values of $A \cup B$ in the previous states
of $\sigma'$.  \qed
%A \PLTL{} property depends in a state only on the prefix of the model that precedes and includes that state.

\begin{theorem}
\label{theo1}
The expressive power of 
\ELTL{} is the same as \QLTL{}.
\end{theorem}
\noindent {\bf Sketch of proof.}
Each \ELTL{} formula $\eta$, as defined in~(\ref{ELTL}),
is expressible using the following equivalent \QLTL{} formula:
\[   \Exists a_1 \ldots \Exists a_n  ( \psi  \wedge   \Box \bigwedge_{1 \leq j \leq n} ( a_j \leftrightarrow \varphi_j))
\]
%\klaus{How do we know that the ``other direction'' below is possible?}
\noindent
For the other direction, one can  first translate the \QLTL{} property  into
a second-order monadic logic formula, then to a deterministic Muller automata and then construct an \ELTL{} formula that holds for the accepting executions of this automaton. The rules of this formula encode the automata states, and the statement describes the acceptance condition of the Muller automaton.
Details are provided in an appendix. \qed

%\EPLTL{}, where the Muller automaton can be
%replaced with a B\"{u}chi automaton,
%where all of its states, except one, sink
%state, are accepting. \qed

\iffalse
The use of the auxiliary propositions $a_1 \ldots a_n$ and the
rules of an \ELTL{} specification 
can be viewed as representing a
state machine, in particular when the only modality that is used in the rules part is $\ominus$.
In that case, we can consider an \ELTL{} specification $\eta$ as consisting of two synchronous parts: an LTL property 
(the statement) 
and a finite automaton, whose states correspond to combinations
of truth values of the auxiliary propositions, and can be referred to
within $\psi$. As a corollary of Theorem~\ref{theo1}, 
this restricted representation is also equivalent to \QLTL{}.
This representation can be useful, for RV, as explored further in this paper, because of the possibility to
incrementally calculate the values of the auxiliary propositions. 
Similarly, it can be
used for extending the expressiveness of LTL model checking.

%\klaus{First sentence below sounds wrong: (1) \PLTL{} already does not have future time operators, and (2) why mention infinite sequences?}
\fi

We define \EPLTL{} by disallowing the future temporal operators (except an implied prefixing $\Box$) in \ELTL{}. This results in
a formalism that is equivalent to a B\"{u}chi automata, where all the states except one are accepting and where the non-accepting state is a sink. 
We can use a related, but simpler construction than in Theorem~\ref{theo1} to prove the following:

\iffalse
In this case, 
we replace the first part of the construction in the proof of Theorem~\ref{theo1}
with the simpler formula $\vee_{s \in A} a_s$, where all the states but one are
accepting.$A$ are the accepting states of the finite automaton.

\klaus{Last sentence above: perhaps say with words that we are always in an acceptance state - if that's what the point is.}
\fi

%For representing safety properties that can be expressed in \QPLTL{}, we can use a restricted version as shown in the following simple Lemma.

\begin{lemma}
\label{simplecase}
The expressive power of 
\EPLTL{} is the same as \QPLTL{}.
%Any safety \QPLTL{} property can be expressed using the following restricted
%\ELTL{} form:
%\[  \Box p {\ where\ } a_j  = \varphi_j \, : \, 
% {j \in \{1, \ldots , n\}}  \] 
% with $p$ being one of the auxiliary variables %$a_j$.
\end{lemma}

\begin{lemma}
\EPLTL{} can with no loss of expressive power be restricted to the form:
\[  \Box p {\ where\ } a_j  = \varphi_j \, : \, 
{j \in \{1, \ldots , n\}}  \] 
with $p$ being one of the auxiliary propositions $a_j$
and $\varphi_j$ contains only a single
occurrence of the $\ominus$ temporal operator (and the Boolean operators). 
\end{lemma}
In this form, the value of the Boolean variables $a_j$ encodes the states of an automaton, and the rules encode the transitions.
%
Yet, this may not be the preferred form
of an \EPLTL{} property, and the use of further past-time temporal operators, and a non trivial statement  part may provide a much more intuitive specification.
%
%The relative expressive power between the propositional temporal logics discussed 
%in this section is summarized in Figure~\ref{prop}. An arrow from A to B means 
%that logic  B is more expressive than A. A double headed arrow shows same 
%expressiveness

\iffalse
\noindent {\bf Proof.} The proof of the Lemma follows from the
axiomatization of \QLTL{} in~\cite{MP}, e.g.,
$\varphi\ \since\ \psi \leftrightarrow \psi \vee ( \varphi \wedge (\varphi\ \since\ \psi))$. \qed

\klaus{In the axiom above, isn't there missing a $\ominus$ before
the second $(\varphi\ \since\ \psi)$?}

\klaus{Above: is the following really correct?: 
``with $p$ being one of the auxiliary variables $a_j$''.}

While this Lemma makes the use of the temporal operators
besides $\ominus$ in \EPLTL{} redundant, it is still
convenient to use a combination of
temporal operators when possible, and
auxiliary variables definitions.
\fi

%\begin{figure}
%\begin{center}
%\includegraphics[width=2.8in]{PROP.jpg}
%\caption{\label{prop}
%Comparing the expressive power of the temporal logics in Section~\ref{propLTL}
%}
%\end{center}
%\end{figure}

\section{RV for Propositional Past-Time LTL and its Extension}
\label{LTLruntime}

Runtime verification of temporal specifications often
concentrates on the past portion of the logic.  Past-time  specifications have the important property that one can distinguish when they are violated after observing a finite prefix of
an execution. For an extended discussion of this issue of {\em monitorability}, see e.g.,~\cite{Ugly,FFM}.

The RV algorithm for \PLTL{}, presented in \cite{HR},
is based on the observation that the semantics of the 
past-time formulas $\ominus \varphi$ and $(\varphi\, \since \,\psi)$ in the current state $i$ is defined in terms of the semantics of its subformula(s)
in the previous state $i - 1$.
To demonstrate this, we rewrite here the semantic definition of the 
%previous-time
%operator $\ominus$, and rewrite the definition
 $\since$ operator in an equivalent form that is
more directly applicable for runtime verification.
\begin{itemize}
%\item $( \sigma , i) \models \ominus \varphi$ iff $i > 1$ and $(\sigma, i-1) \models \varphi$.
\item $(\sigma , i) \models (\varphi \, \since \, \psi)$ iff $(\sigma , i) \models \psi$ or: $i>1$ and
$( \sigma , i)  \models \varphi$ and 
$(\sigma , i-1 ) \models ( \varphi \, \since \, \psi )$.
\end{itemize}
The semantic definition is recursive in both the length of the prefix and the
structure of the property. Thus, subformulas
are evaluated based on smaller subformulas, and the
evaluation of subformulas in the previous state.
The algorithm, shown below, uses two vectors of values indexed by subformulas:  $\old$, which summarizes the truth values of the
subformulas for the execution prefix that
ends just {\em before} the current state, and $\current$, for the execution prefix that
ends with the current state. The order of calculating $\current$ for subformulas is bottom up, according to the syntax tree.
\begin{enumerate}
\item Initially, for each subformula $\varphi$
of $\eta$,
%$B ( \varphi , \current ) = \bfalse $.
$\current ( \varphi ) := \false$.

\item Observe a new event (as a set of propositions) $s$ as input. 
\item Let $\old := \current$.
\item Make the following updates for each subformula. If $\varphi$ is
      a subformula of $\psi$ then $\current ( \varphi )$ is updated before 
      $\current ( \psi )$.
\begin{itemize}
  \item $\current ( \true ) := \true$.
  \item $\current (  \varphi \wedge \psi  ) := 
  \current ( \varphi )\  and\ \current ( \psi )$.
  \item $\current ( \neg \varphi  ) := not\ \current ( \varphi )$.
  \item $\current (  \varphi \since \psi  ) :=  
  \current ( \psi  )\ or\ ( \current ( \varphi ) \ and\ 
      \old ( ( \varphi \since  \psi )))$.
%DAP: I removed the \; around \since, since it penetrated the 2nd column!
  \item $\current ( \ominus \; \varphi ) := \old ( \varphi )$.
\end{itemize}
\item If $\current (\eta) = \false$ then
report a violation, otherwise goto step~2.
\end{enumerate}

\iffalse
\subsubsection*{Extending~\cite{HR}: Runtime verification for \QPLTL{}.}  This can be performed by translating a
formula into a deterministic finite automaton, where one distinguished state, $s$, is a sink state.
Then the sequence of states to be checked forms the input for this automaton, and violation is
announced if $s$ is reached.
\QPLTL{} can be extremely compact, in particular when using alternating existential and universal 
quantifiers, which can result in a huge (non-elementary) explosion of
the automaton representing the same
specification~\cite{Thomas}.
\fi

\noindent {\bf Runtime verification for \EPLTL{}.} For \EPLTL{},  we need to  add to the above
algorithm calculations 
of $\current ( a_j )$ and
$\current ( \varphi_j )$ for each rule of
the form $a_j := \varphi_j$ (the corresponding
$\old$ entries will be updated as in line~3 in the above algorithm).
%
%(calculations
%for the statement part $\psi$ remain the same).
%$\current ( a_j )$ and $\old ( a_j )$,
%and perform the following between steps~3 and~4 %for each $a_j$: $\current ( a_j ) := \current ( %\varphi_j )$.
%
%For \PLTL{}, we can calculate  $\current$, for subformulas,
%bottom up. But
Because the
auxiliary propositions can appear recursively in \EPLTL{} rules, the order of calculation is
subtle. To see this, consider, for example, Formula~(\ref{form3}). It contains
the definition $q := \ominus \neg q$. We cannot calculate this
bottom up, as we did for \PLTL{}, since $\current ( q )$ is not computed yet, and we need 
to calculate $\current ( \ominus \neg q )$ in order 
to compute $\current ( q )$.
However, notice that the calculation is not dependent on the
value of $q$ to calculate $\ominus \neg q$; 
in Step~4 above, we have that
$\current ( \ominus \; \varphi ) := \old ( \varphi )$
so  $\current ( \ominus \neg q ) := \old ( \neg q )$.
%

{\em Mixed evaluation order.}
Under mixed evaluation order, one calculates $\current$ as part of Step~4 of the above algorithm in the following order. 
%The first~four steps, $a$-$d$, are for the rules of the specification $\eta$, and the fifth step $e$ is for the statement $\psi$ of  $\eta$.
\begin{description}
\item{\em a.} Calculate 
$\current ( \delta )$ for each subformula $\delta$
that appears in $\varphi_j$ of a rule
$a_j := \varphi_j$, but {\em not} within the scope of
a $\ominus$ operator (observe that
$\current ( \ominus \gamma )$ is set to $\old ( \gamma )$).
\item{\em b.} Set 
$\current ( a_j )$ to $\current ( \varphi_j )$ for each $j$.
\item{\em c.} Calculate $\current ( \delta )$ for each subformula $\delta$
that appears in $\varphi_j$ of a rule
$a_j := \varphi_j$ {\em  within} the scope of
a $\ominus$ operator.
\item{\em d.} Calculate $\current ( \delta )$ for each subformula $\delta$
that appears in the statement $\psi$, using the calculated $\current ( a_j )$.
%\item{\em e.} Calculate $\current ( \psi )$ for the statement
%$\psi$.
\end{description}

\section{First-Order LTL}

%section*{Syntax and Semantics}
\label{sec:syntax-semantics}

Assume a finite set of infinite domains\footnote{Finite domains are handled with some minor changes, see~\cite{HPU}.}
$D_1, D_2, \ldots$,
e.g., integers or strings. 
Let $V$ be a finite set of {\em variables}, 
with typical instances $x$, $y$, $z$.
%We denote by $x: D$ the fact that the variable $x$ has the domain $D$.
An {\em assignment} over a set of variables $V$
maps each variable $x \in V$ to a value from
its associated domain $\domain ( x )$, where multiple variables (or all of them)
can be related to the same domain. For example
$[ x \rightarrow 5 , y \rightarrow \text{``abc''} ]$ assigns
the values $5$ to $x$ and the value ``abc'' to $y$.
%Let ${\cal R}$ be a set of {\em relations} 
%with typical instances $R$.
%Each relation is associated with some position-dependent domains.

%\klaus{Above: let T be ... Is T used later? The relations are used in the next paragraph, why not introduce them there? Also, relations are not defined really.}
%\klaus{I don't quite understand the motivation below for this way of modeling events: relations with a time stamp, and the whole story of the time stamp not being mentioned etc. Why not just the usual models: traces? Too late to change now for CAV though.}

We define models for \FLTL{} based on {\em temporal} relations~\cite{Chomicki}, that is, relations with
last parameter
that is a natural number, representing a time instance
in the execution. So a tuple of
a relation $R$ can be $(``a", 5, ``cbb", 3)$, where $3$ is
the value of the time parameter. The last parameter $i$ represents a discrete progress of time
rather than modeling {\em physical real time}. It
is used to allow the relations to have
different tuples in different instances of $i$,
corresponding to states in the propositional temporal
logics.

For a relation $R$, $R [ i ]$ is the relation obtained from $R$ by
restricting it to the value $i$ in the last parameter, and removing that last $i$ from the tuples. For simplicity, we will describe henceforth
the logic with relations $R$ that have exactly two parameters, the second of which is the time instance. 
Hence $R [ i ]$ is a relation with just one parameter over a domain
that will be denoted as $dom ( R )$. The definition of the logic over other numbers of parameters is quite straightforward. Our implementation of runtime verification, and the examples described later, fully support relations with zero or more parameter. 
%This also applies to the extensions of the logic in the rest of this section, and to some of the examples used.

%\klaus{Above, what about instead of writing: ``this domain is fixed over all the values of the second parameter $i$'' to write:
%``this domain is independent of the values of the second %parameter $i$'' - if that is what you mean to say.}

%\klaus{Above: we need to say explicitly what $i$ intuitively is meant to represent: positions, time, ....}

\noindent{\bf Syntax.} 
The formulas of the core \FLTL{} logic are 
defined by the following grammar, where $p$ denotes a relation,
$a$ denotes a constant and $x$ denotes a variable. 
\begin{center}
$\varphi ::= \true  \; | \;
    p ( a ) \; | \;
    p ( x ) \; | \;
    ( \varphi \wedge \varphi ) \;  |   \;
   \neg \varphi \; | \;
   \bigcirc \varphi \; | \; 
   ( \varphi \; \until \; \varphi ) \; | $ %\\
   $ \ominus \; \varphi \; | \;
    ( \varphi  \; \since  \; \varphi ) \; | \;
    \exists x \; \varphi$
\end{center}

%We interpret a predicate $p(a)$ as a subformula that
%represents the occurrence of some
%For example, if we open
%a file named $\text{``xyz''}$, we may have $\mathit{open}
%(\text{``xyz''})$, where the
%domain of the predicate ${\mathit{open}}$ is over strings representing
%file names.

\noindent 
Additional operators are defined as in the propositional logic. We define
$\forall x \; \varphi = \neg \exists x \neg \varphi$.
Restricting the modal operators to the past operators
($\since$, $\ominus$ and the ones derived from them) 
forms the logic \PFLTL{}.

\noindent{\bf Semantics.}
A model is a set of temporal
 relations ${\cal R} = \{ R_1 \, \ldots , R_m \}$.
 Since the standard definition of temporal logic is
 over a sequence (``the execution''), let
 ${\cal R} [ i ] = \{ R_1 [ i ] \, \ldots , R_m [ i ] \}$. ${\cal R} [i]$ represents a {\em state}.
 A model ${\cal R}$ can thus be seen as a sequence
 of states ${\cal R}[1] {\cal R}[2] \ldots$. Let 
 $m$ be a bijection
from relation names (syntax)
onto the relations ${\cal R}$ (semantics).
%After the formal semantical definition, we will abuse notation
%and use the same symbol for the representation of the relation in the logic and for the relation as mathematical object.

%\klaus{Above, did you not already say the following before:
%``we denote by $R_k [ i ]$ the restriction of
% the relation $R_k$ to the value $i$ of the last parameter.''.}

\iffalse
At a given state the formula
$p(\text{``a''})$ means that $p (\text{``a''} )$ holds
in the current state,
that is, $p (\text{``a''} )$ is among 
the ground predicates of the state.
Consider now the formula $p ( x )$, for a variable $x \in V$.
We interpret it such that $x$ is assigned any value ``$a$'' where
$p ( \text{``a''} )$ appears in the current state. 
%DP: Added parentheses on the outside level.
Thus, for interpreting $(p ( x ) \wedge q ( y ))$ in a state that
has the predicates
$p ( \text{``a''} )$ and $q ( 3 )$,
we have the assignment $[ x \mapsto \text{``a''} , y \mapsto 3 ]$.
%The past-time temporal operators have the following intuitive meaning.
The formula $(\varphi_1\; \since\; \varphi_2)$ 
(reads $\varphi_1$ {\em since} $\varphi_2$)
means that $\varphi_2$ occurred in the past (including now)
and since then (beyond that state) $\varphi_1$ has been true. This is the 
past dual of the commonly used %DP: changed "common" to "commonly used".
future time  {\em until} modality~\cite{MP}. 
The property $\ominus \; \varphi$ means that $\varphi$ is true 
in the previous state.
This is the past dual of the %DP: removed "common".
future time {\em next} modality.
We can also define the following additional temporal operators:
$P \; \varphi = (\true \, \since \, \varphi)$ (``previously''),
and $H \varphi = \neg P \neg \varphi$ (``always in the past'' or ``historically'').
The operator $[\varphi_1,\varphi_2)$, borrowed from \cite{MaC}, 
has the same meaning as $(\neg \varphi_2\; \since\; \varphi_1)$, but reads more naturally as
a semi-open interval.

%Consider now the formula $P \neg p ( v )$, meaning that
%sometimes in the past $\neg p ( v )$. The semantics of $P$
%will be the union on the sets of values in the previous sets,
%thus, we need to take the union of $\neg p ( v )$ interpreted
%over all past states (including the current one).
%
%Suppose that we have the predicate $p ( i )$ at state $s_i$.
%There are several ways to interpret that.
%\begin{enumerate}
%\item We complement with respect to
%all the values in $domain ( p )$.
%Thus, the complementation will be $N \setminus \{ i \}$ at state $i$,
%where $N$ is the domain natural numbers, and the 
%$N \setminus \{ i \}$ from $1$ to $n$ is $N$.
%
%\item We complement with respect to
%the values in $domain ( p )$ that appeared so far in the sequence.
%Thus, at state $s_i$ we already saw $\{1 , \ldots , i \}$
%and as we see $p ( i )$ in the current state, the complementation 
%at state $s_i$ is $\{ 1, \ldots , i-1 \}$. The union 
%up to the $n$th state will give $\{1, \ldots , n-1 \}$.
%
%\item When new values appear later in the sequence, we 
%update the complementation related to previous states to include
%newly appearing values.
%Then we have to take the union of $n$ sets
%that do not include one distinct element each. This
%gives us at the $n$th state $\{ 1 , \ldots , n \}$.
%\end{enumerate}

%In our implementation we will take the first interpretation.
%However, we will assume that we have a reasonable
%limit on the number of different values that can appear in
%each domain. As it will cost us very little to assume quite
%a large size of values (the size of memory and amount of
%time of the implementation will be propositional to the
%{\em actual} number of values seen and not to this limit), we deem
%this assumption quite reasonable. %We will explain how to 
%remove this constraint (however, this requires quite a
%complicated implementation). It is also quite easy to see
%how to change the semantics to conform with case $1$ and
%how to implement it.

Let $\gamma$ be an assignment to the variables that appear
free in a formula $\varphi$.
Then $( \gamma ,  \sigma, i ) \models \varphi$
if $\varphi$ holds for the prefix $s_1 s_2 \ldots s_i$ of 
the trace $\sigma$
with the assignment $\gamma$. This is a standard definition,
agreeing, e.g., with~\cite{Basin}. Note that by using past %DP: appearing in, instead of appearing with.
operators, the semantics is not affected by
states $s_j$ for $j>i$.
%
\fi

Let $\vars ( \varphi )$ be the set of free (i.e., unquantified) variables of a
subformula $\varphi$. 
%The interpretation of a subformula
%$\varphi$ over some finite sequence is as a set of assignments
%to the variables $\vars ( \varphi)$.
We denote by $\gamma |_{\vars ( \varphi )}$ the restriction (projection) of
an assignment $\gamma$ to the free variables appearing in $\varphi$.
%We are careful in combining the sets of assignments
%between different
%subformulas, e.g., through union and intersection (corresponding
%to disjunction and conjunction, respectively) and the quantification.
%This is done using the projection and extension functions over
%assignments defined above.
%
Let $\epsilon$ be the empty assignment (with no variables). 
%
In any of the following cases, $( \gamma,  {\cal R} , i ) \models \varphi$
is defined where $\gamma$ is an
assignment over $\vars ( \varphi )$, and $i\ge 1$.

\begin{itemize}

\item $( \epsilon , {\cal R} , i ) \models \true$.

\item $( \epsilon ,  {\cal R} , i )
 \models p ( a ) $ if $m(p) ( a, i )$, where $a$ denotes a constant from  $dom (m(p))$.

\item $( [ x \mapsto a ] , {\cal R} , i ) \models p ( x )$ if $m(p)  ( a , i )$, where $\domain (x) = dom ( m ( p ))$.

\item $( \gamma,  {\cal R} , i ) \models ( \varphi \wedge \psi ) $~if~$(
\gamma |_{\vars  ( \varphi )} , {\cal R} , i ) \models \varphi$~and 
$( \gamma |_{\vars ( \psi ) } , {\cal R} , i ) \models \psi$.

\item $( \gamma , {\cal R} , i ) \models \neg \varphi$ if not $( \gamma , {\cal R} , i ) 
\models \varphi$.

\item $( \gamma , {\cal R} , i) \models \bigcirc \varphi$ if 
$(\gamma , {\cal R} , i + 1) \models
\varphi$.

\item $( \gamma , {\cal R} , i ) \models ( \varphi \; \until \; \psi )$ if
for some $j$, $j \geq i$, $( \gamma |_{\vars ( \psi )} , {\cal R} , j) 
\models \psi $ and for each $k$,  $i \leq k < j $,
$( \gamma |_{\vars ( \varphi )} , {\cal R} , k) \models \varphi$.

\item $( \gamma , {\cal R} , i) \models \ominus \varphi$ if $i>1$ and
$(\gamma , {\cal R} , i-1) \models
\varphi$.

\item $( \gamma , {\cal R} , i ) \models ( \varphi \; \since \; \psi )$ if
for some $j$, $1 \le j \leq i$, $( \gamma |_{\vars ( \psi )} , {\cal R} , j) 
\models \psi $ and for each $k$, $j < k \leq i$,
$( \gamma |_{\vars ( \varphi )} , {\cal R} , k) \models \varphi$.

\item $( \gamma , {\cal R} , i) \models \exists x \; \varphi$ if
there exists $a \in \domain ( x )$ such that\footnote{$\gamma \, [ x \mapsto a ]$
is the overriding of $\gamma$ with the binding $[ x \mapsto a ]$.}
$( \gamma \, [ x \mapsto a ], \sigma , i) \models \varphi$.

\end{itemize}

\iffalse
The definition of the {\em since} operator $S$ can be simplified in a standard
way such that it refers only to the positions $i$ and $i-1$ in
the sequence $\sigma$. This is based on the
fact that according to the semantics of {\em since},
$(\varphi \, \since \, \psi) = ( \psi \vee (\varphi \wedge \ominus 
(\varphi \, \since \, \psi )))$.
This will serve in the
implementation to work with only two versions of
the sets of assignments, for the current and previous state:

\begin{itemize}
\item $( \gamma , {\cal R} , i ) \models ( \varphi \, \since \, \psi )$ if
$( \gamma |_{\vars ( \psi )} , {\cal R} , i ) \models \psi$ or $i>1$,
$( \gamma |_{\vars ( \varphi )} , {\cal R} , i ) \models \varphi$, and \\
$( \gamma , {\cal R} , i-1 ) \models ( \varphi \, \since \, \psi )$.
\end{itemize}
\fi

%We can also define the following useful operators:
%$P \varphi = (\true \, \since \, \varphi)$ (for ``previously''),
%$( \varphi \, R \, \psi ) = \neg ( \neg \varphi \, \since \, \neg \psi )$ 
%(the dual of the Since operator), and
%$H \varphi = ( \false \, R \, \varphi )$ (for ``always in the past'').

%We define $\mathbf{F}$ and $\mathbf{T}$ as special constants
%in order to interpret formulas without free variables.
%They are the complement of each other.
%When combined with set operators (complementation, union, intersection),
%$\mathbf{F}$ behaves as the empty set, hence is idempotent to
%union. Accordingly, $\mathbf{T}$ is idempotent to intersection.

\iffalse
We denote by $A_{\vars ( \varphi )}$ the set of all possible assignments
of values to the variables that appear free
in $\varphi$. Thus,
$I [ \varphi, \sigma , i] \subseteq A_{\vars ( \varphi )}$.
To simplify definitions, we add
a dummy position $0$ for sequence $\sigma$ (which starts with $s_1$), 
where every formula is interpreted as an empty set.
Observe that the value $\emptyset$ and $\{ \epsilon \}$, behave
as the Boolean constants $0$ and $1$, respectively.
The set semantics is defined as follows, where $i \ge 1$.

\begin{itemize}
\item $I [ \varphi , \sigma , 0 ] = \emptyset$.
\item $I [ \true , \sigma , i ] = \{ \epsilon \}$.
\item $I [ p ( a ) , \sigma , i ] =$ if $p ( a ) \in \sigma [ i ]$ then
$\{ \epsilon \}$ else $\emptyset$.
\item $I [ p ( v ) , \sigma , i ] = \{ [ v \mapsto a ] \; | \; p ( a ) \in
\sigma [ i ] \}$.
%\item $I [ \true , \sigma , i ] = D^n$.
%\item $I [ ( \varphi \vee \psi ) , \sigma , i ] = 
%I [ \varphi , \sigma , i ] \cup I [ \psi , \sigma , i ]$.
\item $I [ ( \varphi \wedge \psi ) , \sigma , i ] = 
I [ \varphi , \sigma , i ] \;  \bigcap \; I [ \psi , \sigma , i ]$.
\item $I [ \neg \varphi , \sigma , i ] = 
A_{\vars ( \varphi )} \; \setminus \; I [ \varphi , \sigma , i ]$.
\item $I [ ( \varphi \; \since \; \psi ) , \sigma , i ] = 
I [ \psi , \sigma , i ] \; \bigcup \;
( I [ \varphi , \sigma , i ] \; \bigcap \; 
I [ (\varphi \since \psi ) , \sigma , i - 1 ] )$.
\item $I [ \ominus \varphi , \sigma , i ] = I [ \varphi , \sigma , i-1 ]$.
\item $I [ \exists x \; \varphi , \sigma , i ] = 
\proj ( I [ \varphi , \sigma , i ] , \{ x \} )$.
\end{itemize}

\noindent
As before, the interpretation for the rest of the operators can
be obtained from the above using the connections between the operators,
e.g., $I [ P \, \varphi  , \sigma , i ] = 
I [ ( \true \, \since \, \varphi ) , \sigma , i ]$.
The correspondence between this set based semantics 
and the previous semantics, namely that
$\gamma \in I [ \varphi , \sigma, i ]$ iff
$( \gamma , \sigma , i ) \models \varphi$
can be proved by a simple structural induction on
the size of the formulas.
 \fi

For an \FLTL{} (\PFLTL{}) formula with no free variables, denote ${\cal R} \models \varphi$ when $( \epsilon , {\cal R} , 1 ) \models \varphi$.
We will henceforce, less formally, use the same
symbols both for the relations (semantics) and their representation in
the logic (syntax).
Note that the letters $p, q, r$, which were
used for representing propositions in the propositional versions
of the logic in previous sections, will represent relations in
the first order versions.
The quantification over values of variables, 
denoted with $\exists$ and
$\forall$,
here is {\em static} in the sense that they are independent of the state in the execution. 
%Dynamic quantification, as defined for LTL in Section~\ref{sec:extending-prop-ltl},
%and will extend \FLTL{} below, is expressed with $\Exists$ and $\Forall$.

%\section*{Extending First-Order LTL}

%%this is specification formalism that is very expressive and yet monitorable under run time verification~\cite{HPU}.

We demonstrate that the lack of expressiveness
carries over from LTL (\PLTL{}) to \FLTL{} (\PFLTL{}).

\vspace{1ex}
\noindent {\bf Example 1.}
Let $p$ and $q$ be temporal relations (recall that we omit writing the time parameter $i$, denoting $p ( a, i)$ as
$p(a)$). The
specification that we want to monitor is that for each value
$a$, $p(a)$ appears in all the states where $q(a)$ has appeared an even number of times so far (for the odd occurrences, $p(a)$ can also appear, but does not have to). To show that this is not expressible in \FLTL{} (and \PFLTL{}),
consider models (executions) where only one data element $a$ appears. We can prove, by a structural induction on subformulas,
that in this case, for each property $\varphi$
we can replace occurrences of variables within
predicates by $a$, i.e., $p(x)$ and $q(x)$ become $p(a)$
and $q  (a )$, respectively, and that we can throw away
the quantification ($\forall x$, $\exists x$), obtaining
an equivalent formula. This is then equivalent to
an LTL formula that is obtained by replacing
the occurrences of $p(a)$ and $q(a)$ by propositions
$p_a$ and $p_b$, which reduces to Wolper's example~\cite{Wolper}.
Using parametric automata as a specification formalism, as in~\cite{Grum,havelund-rv-data-2018,Meredith2011,Reger2015}, can express this property, 
where for each value $a$ there is a separate automaton that counts the number of times that $q(a)$ has occurred.

\vspace{1ex}
\noindent {\bf Example 2.}
Consider the property that asserts that when
$\mathit{report}(y , x, d)$ appears
in a state, denoting that  process 
$y$ sends some data $d$ to a process $x$,
there was a chain of states with process spawns
$\mathit{spawn}(x, x_1)$, $\mathit{spawn} (x_1, x_2)$ $\ldots$
$\mathit{spawn}(x_l , y)$. i.,e., $y$ is a descendent process of $x$.
The required property needs to
assert about the transitive closure
of the relation $\mathit{spawn}$. 
\FLTL{} can be translated (in a way similar to the standard translation of $LTL$ into monadic first order logic formula~\cite{Thomas}) to a first order formula, with explicit occurrences of time variables
over the naturals and the linear
order relation $<$ (or $\leq$) between them.
%Here, the relations will be written with their explicit last ``time'' parameter, and the temporal operators are replaced with first order quantification.
For example, $\Box \forall x \, (p(x) \rightarrow \Diamond q(x))$ will be translated into $\forall x \, \forall t \, (p(x, t) \rightarrow  \, \exists t' \, (t \leq t' \wedge q(x, t') ) )$.
%The obtained model has, in addition
%to the aforementioned relations, also
%the linear order relations $\leq$ over the naturals.}.
However, the transitive closure of $\mathit{spawn}$ cannot be expressed in first order setting. This can be
shown based on the
compactness theory of first order logic~\cite{Flum}.

\vspace{1.5ex}
\noindent {\bf Extending \FLTL{} with dynamic
quantification.} Relations play in \FLTL{} a
similar role to propositions in LTL.
Hence, in correspondence with the relation
between LTL and \QLTL{}, we
extend \FLTL{} (\PFLTL{}) with dynamic quantification over relations, obtaining \QFLTL{}
(and the past-restricted version \QPFLTL{}). 
The syntax includes $\Exists p \, \varphi$, 
where $p$ denotes a relation.
We also allow $\Forall p\ \varphi = \neg \Exists p \neg  \varphi$. The semantics is as follows. %We denote by ${\cal R} |_Q$ the set of temporal relations obtained from ${\cal R}$ by removing the set of relations $Q$.

%\klaus{Below, has the notation ${\cal R}' |_{\{ q \} }$ been defined?}

\begin{itemize}
\item $(\gamma ,  {\cal R}, i) \models \Exists q \, \varphi$ iff there exists
${\cal R}'$ such that ${\cal R}' \setminus {\{ q \} } = {\cal R}$ and 
$( \gamma, {\cal R}' , i) \models \varphi$.
\end{itemize}
Consequently, quantification over relations effectively extends the model ${\cal R}$ into a model ${\cal R}'$ within the scope of the quantifier. Note that quantification here is dynamic (as in \QLTL{} and \QPLTL{}) since the relations are temporal and can have different sets of tuples in different states.

\vspace{1.5ex}
\noindent {\rm Extending \FLTL{}  with rules.}
We now extend \FLTL{} into \EFLTL{} in a way that is motivated by the  propositional extension from LTL (\PLTL{}) to \ELTL{} (\EPLTL{}). We allow the following formula:
\begin{equation}
\label{EFLTL}
\ \ \ \ \ \psi \mathit{\ where\ } r_j  ( x_j ) := 
\varphi_j (x_j) : j \in \{ 1 , \ldots , n \} \mathrm{\ \ \ \     such\ that,}
\end{equation}
\begin{enumerate}
\item $\psi$, the {\em statement}, is an \FLTL{} formula with
no free variables, 
\item $\varphi_j$ are \PFLTL{} formulas with a single
free variable $x_j$,
\item $r_j$ is an auxiliary temporal relation with two parameters: the first parameter is of the same 
type as $x_j$ and the second one is, as usual,
a natural number that is omitted in the temporal formulas. 
An auxiliary relation $r_j$ can appear within $\psi$. They can also appear in $\varphi_k$ of a {\em rule} $r_k := \varphi_k$, but only within the
scope
of a previous-time operator $\ominus$. 
\end{enumerate}

\noindent
We define the semantics for the \EFLTL{} (\EPFLTL{})
specification (\ref{EFLTL}) 
by using the following equivalent
\QFLTL{} (\QPFLTL{}, respectively) formula\footnote{ Formal semantics 
can also be given by constructing a set of temporal relations extended
with the auxiliary ones
inductively over growing prefixes, as would be done in a detailed proof of Lemma~\ref{samesame}.}:

\begin{equation} \label{equiv}
\Exists r_1  \, \ldots \, \Exists r_n \,  ( \psi \wedge \Box \bigwedge_{j \in \{ 1, \ldots , n\}} (r_j ( x_j )  \leftrightarrow 
\varphi_i ( x_j) )
\end{equation}

The logic \EPFLTL{} is obtained by restricting the temporal modalities of \EFLTL{} to the past ones:
$\since$ and $\ominus$, and those derived from them.

\iffalse
[Well foundedness of auxiliary propositions] \label{fourone}
The values of the the
auxiliary variables 
of an \ELTL{} formula $\varphi$
are uniquely defined in a state of an execution
by the prefix of the execution up to and including that state.
\fi

\begin{lemma}[Well foundedness of auxiliary relations] \label{samesame}
The auxiliary temporal relations of an \EFLTL{} formula 
at state $i$ are
uniquely defined by the prefix of the execution up to and including that state.
\end{lemma}
\noindent {\bf Proof.} By a simple induction, similar to Lemma~\ref{fourone}. \qed
\vspace{1ex}
The following formula expresses the property described in Example~1, which was shown to be 
not expressible using \FLTL{}.
\begin{equation}
\Box \forall x \, (r(x)\rightarrow p(x)) \mathrm{\ where\ }
r(x) = ( q(x) \leftrightarrow \ominus \neg  r(x)) 
\label{eq:wolper-first-order}
\end{equation}
The property that corresponds to Example~2 appears as the third checked example in the implementation section~\ref{implement}.

Analogously to the propositional case, it is easy to show that \EPFLTL{} does not
need the past temporal operators besides $\ominus$.

%\klaus{Something seems wrong in the above formulation.
%E.g.: Is the $\Box$ a restriction?, what is ``fully quantified single occurrence of some auxiliary relation $r_j$''?}

\ignore{
\noindent {\bf Proof.} Basically, the proof is the same as Lemma~\ref{simplecase}, based on the axiomatization of the
past temporal operators. \qed}

%\klaus{Is the formula above really correct?}

%We showed in Theorem~\ref{theo1} that \ELTL{} (\EPLTL{})
%and \QLTL{} (\QPLTL{}, respectively) have the same
%expressive power. We study now the relationship between
%\EFLTL{} (\EPFLTL{}) and \QFLTL{} (\QPFLTL{}, respectively).

\begin{theorem}
\label{theo2}
The expressive power of \EPFLTL{} is strictly weaker than that of \QPFLTL{}.
\end{theorem}

\noindent {\bf Sketch of Proof.}
The proof of this theorem includes encoding of a property that
observes sets of data elements, where 
elements $a$, appears separately, i.e., one per state, as
$v (a)$,
in between states where $r$ appears. The domain of data elements is unbounded.
%(i.e., $r$ is a relation with
%one parameter: the time parameter $i$).
The set of $a$-values observed in between two consecutive $r$'s is called a {\em data set}.
The property asserts that there are no two consecutive data sets that are equivalent.
This property can be expressed in \QPFLTL{}. The details appear in an appendix.

We use a combinatorial argument to show by contradiction that one cannot express this property using any
\EPFLTL{} formula $\varphi$. The reason is that every prefix of a model for an \EPFLTL{} property is extended uniquely with
auxiliary relations, according to Lemma~\ref{samesame}. Each prefix can be summarized by a finite number of relations: the ones in
the model, the auxiliary relations and the
assignments satisfying the subformulas. The size of
each such relation is bounded by ${\cal O} (  m ^ N )$
where $m$ is the number of values appearing in the
prefix, and $N$ is the number of parameters of the relations.
However, the number of different data sets over $m$ values is
$2^m$. This means that with large enough number of different values, each \EPFLTL{} formula $\varphi$ over the models of this property can have two prefixes with the same summary, where one of them has a data set that the other one does not. The semantics of \EPFLTL{} implies that extending
two prefixes with the same summary in the same way would have the same truth value. Consequently, we can extend
the two prefixes where some data set  appears in one of them
but not  in the other into a complete model, and $\varphi$ will not be able to distinguish between these models. \qed

%The relative expressive power between the first-order %temporal logics presented in this Section
%appear in Figure~\ref{fsfr}.

%It follows from~(\ref{equiv}) that \EFLTL{} (\EPFLTL{}, %respectively) has the same
%expressive power as \QFLTL{} (\EPFLTL{}, respectively) %restricted to having its
%dynamic 
%quantification to being existential, and prefixing %the
%rest of the formula.

\vspace{0.7ex} From Theorem~\ref{theo2} and Equation~(\ref{equiv}) we immediately obtain:
\begin{corollary}
Restricting the quantification of \QPFLTL{} to existential quantification, strictly weakens its expressive power\footnote{It is interesting to note that for~\QPLTL{}, restriction to existential quantification does not change the expressive power.}.
\end{corollary}

\ignore{

\begin{figure}
\begin{center}
%\includegraphics[height=1.5in,width=3in]{FIRST.jpg}
\includegraphics[width=3in]{FIRST.jpg}
\caption{\label{fsfr} An arrow from A to B means that logic  B is more expressive than A}
\end{center}
\end{figure}
}

\section{RV for Past-Time
First-Order  LTL and its Extension}
\label{EPLTLRV}

Runtime verification of \FLTL{} is performed on an
input that consists of {\em events} in the form of
tuples. In our notation, the input
consists of a sequence ${\cal R}[1] \, {\cal R}[2]  \ldots$,
which we earlier identified with states, where each
${\cal R}[i]$ consists of the relations in
${\cal R}$ with
the last parameter is restricted to  $i$. A typical use
of runtime verification restricts the
events (tuples from the relations in ${\cal R}$) for
each state to a finite number, and often even to a single event.
The RV algorithm will also make
use of sets of assignments over a set of variables, for satisfying a subformula at
some state, corresponding to $\old$ and $\current$, also represented as relations
(instead of propositions, as used for LTL in Section~\ref{LTLruntime}).

\ignore{
\klaus{In the text above there are some imprecisions, such as
(1) ``tuples of relations'' - relations are sets of tuples, we don't have tuples of relations. 
(2) ``${\cal R}[i]$ consists of some relations'': it is a set of
relations (I think).
(3) ``to have, per each state just some finite tuples'': tuples are finite. You mean: to contain a finite set of tuples?
(4) what is the story about a set of assignments also being a relation? Needed?
}}

%\subsubsection{Syntax}

\ignore{
\klaus{Below: (\ref{EFLTL}) was not a \EPFLTL{} form of specification.}}

\iffalse
, restricted to the past operators. We now present the definition of \EPFLTL{}, the final main logic proposed in this paper. 
The syntax is as follows:

\[ 
\psi \mathrm{\ where\ } r_j ( x_j) := 
\varphi_j (x_j) : j \in \{ 1 , \ldots , n \}
\]
\begin{center}
$\varphi ::= \true  \; | \;
    p ( a ) \; | \;
    p ( x ) \; | \;
    ( \varphi \wedge \varphi ) \;  |   \;
   \neg \varphi \; | \;
    \ominus \varphi \; | \;
    ( \varphi  \; \since  \; \varphi ) \; | \;
    \exists x \; \varphi$
\end{center}

\klaus{NOte t that $\psi$ above is not defined as a syntactic category. It should really be $\varphi$. Likewise in other parts of the paper. Perhaps too big a change now? Also, there should be
a greek letter for a specification. E.g.:

\[ 
\pi ::= \varphi \mathrm{\ where\ } r_j ( x_j) := 
\varphi_j (x_j) : j \in \{ 1 , \ldots , n \}
\]
\[
\varphi ::= ...
\]}
\fi

\iffalse
\noindent {\bf Set Semantics.}
We now refine the semantics of the logic. Under the new definition, 
$I [ \varphi , \sigma, i ]$ is a function that returns
a set of assignments such that $\gamma \in I [ \varphi , \sigma, i ]$ 
iff $( \gamma , \sigma , i ) \models \varphi$.
This redefinition will later lead to a simple implementation 
using BDDs, where each set of assignments will be represented
as a BDD, and the Boolean operators will correspond directly 
to Boolean operators on BDDs.

In order to deal with subformulas with different sets of free
variables (hence, different domains for assignments), 
we apply a projection and an extension operator to assignments
over a subset of the variables. Let $\Gamma$ be
a set of assignments over the variables $W$, and
$U \subseteq W$.
Then $\proj ( \Gamma , U )$ (for ``projecting {\em out}'' or  ``hiding''
the variables $U$)
is the largest set of assignments over 
$W \setminus U$, 
each agreeing with some
assignment of $\Gamma$ on all the variables in $W \setminus U$.
Let $U \cap W = \emptyset$, then
$\ext ( \Gamma , U )$ is the largest set of assignments
over $W \cup U$, where each such assignment agrees 
with some assignment in $\Gamma$ on the
values assigned to the variables $W$. 
This means that we extend $\Gamma$ by
adding arbitrary values to the variables in $U$ from their domains.
Then $\proj ( \ext ( \Gamma , U ) , U ) = \Gamma$ holds.
We define the union and intersection operators on sets of
assignments, even if they are defined over non identical
sets of variables. 
In this case, the assignments are extended
over the union of the variables. Thus, if $\Gamma$ is a
set of assignments over $W$ and $\Gamma'$ is
a set of assignments over $W'$, then
$\Gamma \; \bigcup \;  \Gamma'$ is defined as $\ext (\Gamma , W' \setminus W ) \cup
\ext ( \Gamma' , W \setminus W' )$ and
$\Gamma \; \bigcap \; \Gamma'$ is $\ext (\Gamma , W' \setminus W ) \cap
\ext ( \Gamma' , W \setminus W' )$.  Hence, both are defined
over the set of variables $W \cup W'$.

%We define $\mathbf{F}$ and $\mathbf{T}$ as special constants
%in order to interpret formulas without free variables.
%They are the complement of each other.
%When combined with set operators (complementation, union, intersection),
%$\mathbf{F}$ behaves as the empty set, hence is idempotent to
%union. Accordingly, $\mathbf{T}$ is idempotent to intersection.

%\klaus{One may of course ask whether the following section on the 
%set semantics is needed. Maybe it is needed, since we probably need a semantics. Something to consider.}
\fi

\vspace{1.4ex} \noindent {\bf Set Semantics.}
The RV algorithm for
(R)\PLTL{}, presented in~Section \ref{LTLruntime}
calculates
$\current ( \varphi )$, for $\varphi$ a
subformula of the monitored property, to be the
truth value of $\varphi$ over the
prefix inspected by the RV algorithm so far. For (R)\PFLTL{},
$\current (\varphi)$ consists of the
sets of assignments (in the form of relations
over the free variables in the subformula),
rather than Boolean variables.

We provide an alternative semantics for 
the logic \EPFLTL{}, without
changing its interpretation, in a
way will be more directly related to the calculation of values
in $\current$ by the
RV algorithm that will be presented below. 
Under the  {\em set semantics}
(introduced in~\cite{HPU} for \PFLTL{}, and extended here for \EPFLTL{}),
$I [ \varphi , \sigma, i ]$ is a function that returns
a set of assignments such that $\gamma \in I [ \varphi , \sigma, i ]$ 
iff $( \gamma , \sigma , i ) \models \varphi$.
We present here only two simple cases of the set semantics. The full set semantics appears in the appendix. 
\begin{itemize}
\item $I [ ( \varphi \wedge \psi ) , \sigma , i ] = 
I [ \varphi , {\cal R} , i ] \;  \bigcap \; I [ \psi , \sigma , i ]$.
\item $I [ ( \varphi \; \since \; \psi ) , {\cal R} , i ] = 
I [ \psi , {\cal R}, i ] \; \bigcup \;
( I [ \varphi , {\cal R} , i ] \ \bigcap \ 
I [ (\varphi \since \psi ) , {\cal R} , i - 1 ] )$.
\end{itemize}

\noindent {\bf Runtime verification algorithm for \PFLTL{}.}
%
We start by describing an algorithm for monitoring 
\PFLTL{} properties, presented in~\cite{HPU} and implemented
in the tool~\dejavu. 
We enumerate data values appearing in
monitored events, as soon as we first see them. We
represent relations over the Boolean
encoding of these enumeration, rather than over the data values themselves. A hash function is used to connect the data values to their enumerations to maintain consistency
between these two representations.
The relations are
represented as BDDs~\cite{Bryant}. 
For example, if the runtime-verifier sees the input 
events 
$\mathit{open}(\text{``a''})$, 
$\mathit{open}(\text{``b''})$, 
$\mathit{open}(\text{``c''})$, 
it will encode them as
$000$, $001$ and $010$ (say, we use 3 bits $b_0$, $b_1$ and $b_2$
to represent each enumeration, with $b_2$ being the most significant bit).
%
A Boolean representation of the {\em set} of values 
$\{\text{``a''},\text{``b''}\} $ would be equivalent to a Boolean function $(\neg b_1 \wedge \neg b_2)$ that returns 1 for $000$ and $001$ ($b_0$ can be arbitrary).

%

Since we want to be able to deal with infinite domains
(where only a finite number of elements may appear in a given observed prefix) and maintain the ability to perform
complementation, unused enumerations represent the
values that have not been seen and their relations
to all other values. 
%The algorithm maintains that these unseen values are correctly represented, as shown in~\cite{HPU}. 
In fact, it is sufficient to have just one enumeration representing these values per each variable of the LTL formula. 
%However, it is important to guarantee that at least one such enumeration exists. 
We guarantee that at least one such enumeration exists by preserving for that purpose the enumeration $11\ldots11$.
%, i.e., the highest Boolean representation,
%to represent values not yet seen.
We present here only the basic algorithm. For versions that
allow extending the number of bits used for enumerations and garbage collection of enumerations, consult~\cite{HP}.

\iffalse
Our preferred representation of a set of values (assignments) is as an Ordered Binary Decision
Diagram (OBDD, although we write simply BDD)~\cite{Bryant}.
A BDD~\cite{Bryant} is essentially a compact representation 
of a Boolean tree, where compaction glues together isomorphic subtrees. Each non-leaf node is labeled with one of the
Boolean variables $b_0,\ldots,b_{k-1}$ (where $k$ is the number of bits used to represent the values of a variable). A non-leaf node $b_i$ 
is the source of two 
arrows leading to other nodes, representing respectively that the node has the value 0 (false) or 1 (true).
%A dotted-line arrow represents that $b_i$ has the Boolean value $0$, while a thick-line 
%arrow represents that it has the value $1$. 
The nodes in the DAG have the
same order along all paths from the root. However, some of the nodes may be
absent along some paths, when the result of the Boolean function does not 
depend on the value of the corresponding Boolean variable. Each path leads 
to a leaf node that is marked by either a $0$ or a $1$, representing the 
Boolean value returned by the function for the Boolean values on the path.
\fi

\iffalse %%%%%
The definition of the {\em since} operator $S$ can be rewritten in a standard
way such that it refers only to the positions $i$ and $i-1$ in
the sequence $\sigma$. 
This is based on the
fact that according to the semantics of {\em since},
$(\varphi \, \since \, \psi) = ( \psi \vee (\varphi \wedge \ominus 
(\varphi \, \since \, \psi )))$.
This will serve in the
implementation to work with only two versions of
the sets of assignments, for the current and previous state:

\begin{itemize}
\item $( \gamma , \sigma , i ) \models ( \varphi \, \since \, \psi )$ if
$( \gamma |_{\vars ( \psi )} , \sigma , i ) \models \psi$ or $i>1$,
$( \gamma |_{\vars ( \varphi )} , \sigma , i ) \models \varphi$, and 
$( \gamma , \sigma , i-1 ) \models ( \varphi \, \since \, \psi )$.
\end{itemize}
\fi %%%%%

%The rest of the operators are defined as follows:
%$\false = \neg \true$, $\forall x \; \varphi = \neg \exists x \neg \varphi$,
%$(\varphi \vee \psi) = \neg ( \neg \varphi \wedge \neg \psi )$.
%We can also define the following useful operators:
%$P \varphi = (\true \, \since \, \varphi)$ (for ``previously''),
%$( \varphi \, R \, \psi ) = \neg ( \neg \varphi \, \since \, \neg \psi )$ 
%(the dual of the Since operator), and
%$H \varphi = ( \false \, R \, \varphi )$ (for ``always in the past'').

%\noindent {\bf Set Semantics.}
%It helps to represent the algorithm to 
%redefine the semantics of
%the logic in terms of sets of assignments satisfying a formula.
%Then $I [ \varphi , \sigma, i ]$ is a function that returns
%a set of assignments such that $\gamma \in I [ \varphi , \sigma, i ]$ 
%%Then a BDD for a set containing the values ``de'' and ``af''
%%(2nd and 3rd values)
%will return $1$ for $01$ and $10$. If the Boolean function is over
%$b_0$ (for most significant bit) and $b_1$ (for least significant), 
%then this is the Boolean function 
%$(\neg b_0 \wedge b_1) \vee (b_0 \wedge \neg b_1)$.

Given some ground predicate $p ( a )$, observed in the monitored execution, 
matching with $p ( x )$ in the monitored property,
let $\lookup ( x, a )$ be the enumeration of $a$ (a lookup in the hash table). If this
is $a$'s first occurrence, then it will be assigned a new enumeration.
Otherwise, $\lookup$ returns the enumeration that $a$ received before. We can use a counter, for each variable $x$, counting the number of different values appearing so far for $x$. When a new value appears, this counter is incremented, and the value is converted to
a Boolean representation. 
%Enumerations that were not yet used represent
%the values not seen yet. 
%In particular, we always leave at least
%one enumeration, $11\ldots 11$, for this purpose.
%In the next section we introduce data reclaiming, which allows 
%reusing enumerations for values that no longer affect the checked 
%property. 
%This involves a more
%complicated enumeration mechanism.
%
The function $\bddset (x, A)$ returns
a BDD that represents the set of assignments where $x$ is mapped to 
(the enumeration of) $v$ for
$v \in A$. This BDD is independent
of the values assigned to any variable
other than $x$, i.e., they can have any value.
For example, assume that we use the three Boolean variables (bits) $x_0$, $x_1$ and $x_2$
for representing enumerations over $x$ (with $x_0$ being the least significant bit), and
assume that $A = \{ a , b \}$, 
$\lookup ( x, a ) = 000$, and $\lookup ( x , b ) = 001$.
Then $\bddset ( x , A )$ is a BDD representation of the Boolean function 
$(\neg x_1 \wedge \neg x_2)$. 
%But the BDD function will be independent of the values of Boolean variables
%representing other variables.
%Notice that $p ( x )$ is
%equivalent, and is represented as $\exists y \, \exists z \, p ( x )$
%when the variables used in the formula are $x$, $y$ and $z$.

Intersection and union of sets of assignments are translated simply
into conjunction and disjunction of their BDD representation,
respectively; complementation
becomes BDD negation. We will denote
the Boolean BDD operators as $\bddwedge$, $\bddvee$ and $\bddneg$.
To implement the existential (universal, respectively) operators, 
%as in the interpretation of $\exists x \; \varphi$, 
we use the BDD existential (universal, respectively) operators over
the Boolean variables that represent (the enumerations of) the values of $x$. 
%Thus, $x$ is translated into
%is represented using the Boolean variables $x_1$, $x_1, \ldots x_{k}$ into 
Thus, if $B_{\varphi}$ is the BDD representing
the assignments satisfying $\varphi$ in
the current state of the monitor, then 
$\bddexists (\langle x_0,\ldots,x_{k-1} \rangle, B_{\varphi})$
is the BDD that represents the assignments satisfying $\exists x \; \varphi$ in the current
state.
Finally, $\bfalse$ and $\btrue$  are the BDDs that return always
$0$ or $1$, respectively. The algorithm is as follows.

%The case where the number of bits used to represent values for
%some domains is eventually insufficient is dealt with by
%expanding the number of bits for the BDD. This will be explained later,
%but is much harder to implement, and may require a lot of overhead
%to update the BDDs while online computation is monitored. This
%may be unnecessarily, as we can assign a large number of bits
%a priory, where these bits represent a set of values that is
%exponentially larger.

\iffalse
The algorithm shown below uses two vectors (arrays) of values indexed by subformulas: 
$\old$ for the state before that event, and
$\current$ for the current
state (after the last seen event).
While in the propositional case, in Section~\ref{LTLruntime}, the vectors contain Boolean values, here
they contain BDDs. The temporal relations,
restricted to each state, are
represented also as BDDs.
The algorithm follows.
\fi

\begin{enumerate}
\item Initially, for each subformula $\varphi$ of $\eta$,
%$B ( \varphi , \current ) = \bfalse$.
$\current ( \varphi ) := \bfalse$.
\item Observe a new state (as a set of ground predicates) $s_i$ as input. 
\item Let $\old := \current$.
\item Make the following updates for each subformula. If $\varphi$ is
      a subformula of $\psi$ then $\current ( \varphi )$ is updated before 
      $\current ( \psi )$.
\begin{itemize}
  \item $\current ( \true ) := \btrue$.
  \item $\current ( p_k ( a ) )$ := if $R_k [ i ] ( a )$ then
  $\btrue$ else $\bfalse$.
  \item $\current ( p_k ( x ) )$ :=
     $\bddset ( x ,  \{ a \ | \ R_k[i] ( a  ) \} )$.
  \item $\current ( ( \varphi \wedge \psi ) ) := 
  \bddwedge ( \current ( \varphi ) , \current ( \psi ))$.
  \item $\current ( \neg \varphi  ) := \bddneg ( \current ( \varphi ))$.
  \item $\current ( ( \varphi\ \since\ \psi ) ) :=  
  \bddvee ( \current ( \psi  ) , \bddwedge ( \current ( \varphi ) ,
      \old ( ( \varphi \ \since \ \psi ))))$.
%DAP: I removed the \; around \since, since it penetrated the 2nd column!
  \item $\current ( \ominus \; \varphi ) := \old ( \varphi )$.
  \item $\current ( \exists x \; \varphi ) := 
  \bddexists (\langle x_0,\ldots,x_{k-1} \rangle, \current ( \varphi ))$.
\end{itemize}
\item If $\current (\eta) = \false$ then
report a violation, otherwise goto step~2.
\end{enumerate}

%The use of BDDs can be replaced by other representations that
%can compactly and efficiently represent sets of values (e.g., ZDDs), and to which one can apply set operations like complementation, intersection, union (which are simply negation, conjunction and disjunction
%in BDDs) and projection (for the quantification).

\noindent {\bf RV algorithm for \EPFLTL{}}
We extend now the algorithm to capture \EPFLTL{}.
The auxiliary relations $r_j$ extend the model, and we need to keep 
BDDs representing $\current ( r_j )$ and
$\old (r_j)$ for each relation $r_j$. We also need to calculate the 
subformulas $\varphi_i$ that appear in
a specification, as part of the runtime verification,
as per the above \PFLTL{} algorithm. One subtle point is that the auxiliary 
relations $r_j$ may be defined in a rule with respect to
a variable $x_j$ as in $r_j (x_j ):= \varphi_j (x_j)$ (this can be generalized 
to any number of variables), but $r_j$ can be used
as a subformula with other parameters in other rules or in the statement e.g., 
as $r_j ( y )$. This can be resolved by a BDD renaming function 
$\rename ( r_j (x_j ) , y)$.
%
We then have to add
following updates to step~4 of the above algorithm, when performing runtime verification of
$\psi$.

\noindent \begin{description}
\item For each rule 
$r_j ( x_j ) := \varphi_j ( x_j )$:
\item{\ \ }  calculate $\current ( \varphi_j )$;
\item{\ \ }  $\current ( r_j )  := \current (\varphi_j)$;
\item{\ \ } $\current ( r_j ( y ) )$ := $\rename ( r_j(x_j) ,  y )$;
\item{\ \ } $\current ( r_j ( a ) )$ := if $\current ( r_j ) ( a )$
then $\btrue$ else $\bfalse$
\end{description}

%\klaus{Above: $\current ( r_j ) ( a )$ has not been defined I think.}

\noindent
As in the propositional case, the evaluation order cannot be
simply top down or bottom up, since relations can appear
both on the left and the right of a definition
such as $r ( x ) := p ( x ) \vee \ominus r ( x )$. 
We need to use the {\em mixed evaluation order}, described in Section~\ref{LTLruntime}.

\vspace{1.5ex}
\noindent {\bf Complexity.} BDDs were first introduced to model checking~\cite{McMillan} since they can often (but not always) allow
a very compact representation of states.
In our context, each BDD in $\old$ or $\current$
represents a relation with $k$ parameters, which summarizes
the value of a subformula of the checked \PFLTL{} or
\EPFLTL{} property
with $k$ free variables
over the prefix observed so
far. Hence, it can grow up to a size that is polynomial in the number of values appearing
in the prefix, and exponential in $k$ (with $k$ being
typically very small). However, the marriage of BDDs and Boolean enumeration is in particular
efficient, since collections of adjacent Boolean enumerations tend to compact well.